Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-04-14 paper.tex
proofs

% Created 2017-04-14 Fri 09:02
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{lmodern}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org2a2b726}

@@

First proposed by \cite{Thorpe2005}, ``passthoughts'' authentication allow users to 
to submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., the unique way that thought is expressed)
in a single step, by performing a single mental task.
Since its original proposal, passthoughts has been validated in lab settings, even with 
consumer-grade devices \cite{Chuang2013b} and in-ear EEG earbuds \cite{curranpassthoughts}.
The protocol appears to be robust against impersonation attacks, validating that it provides two factors of authentication \cite{Johnson2014}.

@@

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life.
First, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughts to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).
Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand 
the space of possible passthoughts,
and how passthoughts change over time (Section 4).

@@

\section{Background}
\label{sec:org246c216}

@@

An ongoing problem in authentication lies in balancing strong security
(i.e., combining multiple factors of authentication)
with usability.
As an example, major industry players such as Google and
Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA), in which a user enters his or her password (a knowledge factor),
and subsequently receives a code on their cellphone (a posession factor).

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgda4a571}

@@

\subsection{Passthought authentication}
\label{sec:org1eb6c99}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org7a20391}

Even consumer-grade headsets can be uncomfortable to wear, and are awkwardly visible to outside observers. Earbuds present a more discreet, comfortable location for an EEG sensor, as many people already wear earbuds in day-to-day life.

@@

\section{User attitudes and perceptions}
\label{sec:org6284e0b}

In spite of past work making passthoughts less visible with more convenient form-factors,
a large question still remains:
What sense would people make of passthoughts, as a technology, in their everyday life?
This question begs not only user-centered design studies with passthoughts itself,
but more general questions about what EEG means to people,
and what people believe EEG data can reveal about them.
Past work has established the almost magical abilities that people tend to ascribe to brain-scanning devices, even subjects with specific training in the limitations of brain-scanners \cite{Ali2014a}.
This section outlines common concerns around ``mind-reading'' machines, and how they relate to EEG and passthoughts specifically.

\subsection{Contending with mind-reading machines}
\label{sec:org605033b}

@@

If people \emph{think} a certain technology measures aspects of mind, it will certainly affect the way they engage with that technology - whether or not it works the way they expect \cite{Ali2014a}. Meanwhile, if they think that a given technology does \emph{not} measure their mind, when it fact it does, users may suffer a breach of what Nissenbaum might call the ``appropriateness of the flow of information'' \cite{Doyle2011}. In both cases, knowing what people expect will help us anticipate their needs, and concerns.

If we wish to understand what role passthought authentication \emph{could} play in day-to-day life,
we must view it both through the lens of potential privacy concerns, \emph{and} through the lens of possible opportunities for self-reflection and self-understanding. 
Of course, users' attitudes will not be fixed: they will evolve over time, as users observe the device in action, and correlate its judgments with their own lived experiences \cite{Nafus2016}.
In the next section, we turn a discussion on EEG specifically as a way of motivating questions around the meaning people may build around passthought authenticators.

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org1129bb0}

@@

\label{fig:rank}
\begin{figure*}
\centering
\includegraphics[width=.9\linewidth]{./figures/rankings.png}
\caption{``Please rank the following sensors in how likely you believe they are to reveal what a person is thinking and feeling.'' Higher bars indicate higher rank, or more likely to reveal what one is thinking or feeling.''}
\end{figure*}

\section{Diversity and security of passthoughts}
\label{sec:orgdd2553c}

@@

While past work on passthoughts has achieved excellent results on corpora of recordings from different users, 
these studies do not consider passthoughts from a variety of different subject conditions.
Passthoughts studies must collect larger, and more diverse corpora of EEG data to examine how EEG signals change (or remain stable) throughout the dynamic contexts of daily life.
For example, sitting subjects may have different patterns of neural activity from subjects who are standing, walking or exercising \cite{Thibault2016a},
let alone subjects who are under the influence of e.g. caffiene or alcohol.

Investigating this question could also help us understand how and why passthoughts work at all: Why are passthoughts unique, and how unique are they?
A primary question in passthoughts surrounds how large the real space of possible passthoughts might be \cite{Thorpe2005}.
While the space of possible passthoughts is potentially unlimited, we do not yet know what passthoughts we stand a reasonable chance at observing consistently over time.
A larger corpus of data might help shed light on this issue by allowing us to observe the distribution of signals that people produce over time.

A more subtle, but closely related question, surrounds how passthought EEG recordings relate statistically to non-passthought EEG recordings.
In other words, we do not know how the particular passthoughts observed in past work are drawn from the distribution of EEG signals that an individual produces over the course of their day.
This blind-spot poses a possible challenge to passthought's vulnerability to dictionary-style cracking.
If an attacker has a large enough corpus of EEG readings, do some passthoughts start to look as guessable as \emph{password1234}?
By answering such questions, we could design data-driven policies for, e.g., how many retry attempts passthought authenticators should allow.

\section{Two studies on passthoughts}
\label{sec:org78fbfbd}

The prior two sections raise two main topics surrounding passthoughts that future work could address. 
First, our limited understanding of passthoughts' usability, and user attitudes about the sensing modality present immediate questions for further development of this technology.
Meanwhile, our limited knowledge of how passthoughts shift and change over time, and around the diversity of EEG signals as our limited (non-medical) devices sense them,
raise questions about how frequently passthoughts would need to be calibrated, how accurate we can expect the protocol to be in different context, and how secure it might remain under threat from a motivated attacker.

This section proposes two studies on passthought authentication which, taken together, could make headway on these topics.
One study, a controlled, lab-based experiment, seeks to raise fundamental questions about how the feedback of a real-time authentication system may affect the performance of passthoughts.
It also begins to address certain, limited questions around the shifting nature of neural signals.
The second study, a longitudinal deployment, seeks to collect a large and diverse corpus of EEG signals, while probing people's beliefs and attitudes about EEG and brainscanning in everyday life.
Together, these studies address both long-term concerns about user attitudes and signal diversities, and also short-timescale questions about the usability and accuracy of passthoughts in realistic use scenarios.

\subsection{A real-time passthought authenticator}
\label{sec:org08c501f}

Passthoughts is supposed to be a more usable form of multi-factor authentication
as compared to existing protocols,
as it provides both a knowledge and an inherence factor in a single step.
However, no study yet has systematically evaluated passthoughts' usability.
Here, we propose a study aimed at examining passthoughts' usability in an ecologically valid context.

\subsubsection{Study protocol}
\label{sec:org1b0cbf0}

@@

After this task, subjects could take a post-questionnaire including various usability questions.
After filling out this questionnaire, the experimenter might engage users in a brief, ten-minute semi-structured interview,
in which subjects were asked to recount their experience with the authenticator, in their own words.
This interview could help gain some richer, qualitative data that traditional survey methods might fail to capture.

\subsubsection{The effect of feedback}
\label{sec:org6038bfe}

@@

Furthermore, using the data collected during this study, we could perform an offline analysis 
to test for the effect of these conditions on the actual performance of users' passthoughts, and consequential classifier accuracy.
When subjects are continuously rejected, do their passthoughts change in frustration (or in an attempt to gain access)?
We might find that passthought performance 
remains stable, regardless of what feedback subjects are shown.
Alternatively, we might find that performance changes 
when subjects are continuously rejected from their authenticator.
Alternatively, performance may change, 
even when subjects are continually accepted by their classifier.

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgd2b8171}

@@

By comparing the final FAR and FRR for each subject using these strategies, 
compared to the one-time calibration strategy, we could begin to get an idea as to whether
this strategy helps achieve superior performance, especially when signals change.
This analysis could also act as a harbinger for some of the possible downsides of this approach:
If a user is continually rejected, and the classifier is re-trained using those rejections as negative examples,

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orga7f95ff}

@@

Unfortunately, these challenges - particularly those around shifting neural signals - also make it difficult to produce a passthought authenticator that works with any reliability in real contexts.
This makes a longitudinal study with a working authenticator impractical for now.
However, we may still perform a longitudinal study that allows us to interrogate the usability aspects around (and attitudes about) passthoughts specifically, and EEG generally.
In so doing, we may also collect a larger and more diverse corpus of passthoughts, which can be used to address the paucity of data we face today.
A technology probe or diary study \cite{Gaver1999} could help address both of these issues at the same time.

@@

\subsubsection{Study protocol}
\label{sec:org6624a8b}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

Such a study would trade a large population size for a large corpus of diverse data.
This tradeoff allows us to closely investigate the diversity of EEG signals within subjects.
The diverse readings encountered in day-to-day life could help us understand how such signals change as a function of time, and/or in different psychophysical states.
At the same time, our small sample size could enable a rich, qualitative understanding of users attitudes.

\subsubsection{A more diverse corpus}
\label{sec:org1374951}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orga1e7779}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

Leveraging the statistical clusters of EEG data generated by these algorithms, it might also be possible to generate a ``passthoughts cracker,'' capable of generating plausible passthoughts. 
Feeding these algorithms into pre-trained passthought classifiers, we can begin to generate realistic models of classifiers' resistance to cracking attempts. 
These cracking experiments could lead to defenses against cracking attempts, by enforcing retry attempt timeouts or other methods for limiting break-in risk, such that strong guarantees can be enforced.

\subsubsection{Usability and attitudes}
\label{sec:org93eda86}

By deploying a real sensing apparatus, be it a traditional consumer device such as the Muse \cite{Mihajlovic2015} 
or a more experimental piece of equipment such as an earbud,
and having people record EEG data in their daily life, we could learn more about the interpretative qualities of these data \cite{NafusDawn;Sherman2014}.
Such a study presents a dual opportunity to understand user beliefs in a rich, qualitative sense, while simultaneously collecting the large, diverse and longitudinal corpus of EEG signals necessary if we wish to stand a chance at decent classification accuracy in the wild.

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:orgee42247}

After the studies described above, 
we will have a much better grasp on the usability, and security properties of passthought authentication.
However, there may still be unexplored risks, challenges, and tradeoffs,
especially around user privacy.
Indeed, some of these risks are unique to the application context, and to EEG as a class of biosignal. 
This section briefly reviews risks to user privacy and security that widespread passthought authentication may introduce. 
We present broad class of categories from which such risks may emerge.

\subsection{Privacy}
\label{sec:org0ce2819}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org239dd78}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

There is also the question of the security of data infrastructures in which EEG data might be stored.
Large data repositories are what Wolf \cite{Wolf2010} calls a ``toxic asset''; they must be maintained, 
lest the maintainer take liability for the potentially harmful fallout of poor data management.
With biosignals, as with many kinds of data, it is not entirely clear what they might mean until they are already collected in aggregate. 
At this point, it is too late to decide on an appropriate data security policy.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orga38c9fb}

@@

\section{Further Future Directions}
\label{sec:org2a60ef2}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:org5aebc34}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to preempt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:org8f2e818}

After immediate challenges are overcome,
one further-out, though potentially exciting possibility is that of using EEG for \emph{continuous authentication}.
Continuous authentication schemes seek to authenticate a user using ongoing streams of data or activity, sometimes by giving a probability that a person's identity is authentic \cite{Bojinov2012}.
Such schemes are a natural match for wearables, which can continuously collect and process biometric data.
A recent startup, Unify.ID, has begun to perform cross-device continuous authentication as a service \cite{UnifyID2017};
however, as a knowledge factor, it currently falls back on traditional passwords, which come with both well-known risks and annoyances to usability.

A continuous passthought authenticator could incorporate both knowledge and inherence factors (along with, optionally, the posession factor of a unique sensing device).
Subjects could perform secret passthoughts for certain unlocking actions,
while the authenticator could fall back on inherence in the base case (e.g. as an additional check on sites where the user's logged-in session would otherwise be remembered).
In theory, this strategy provides better security properties than saved sessions or cookies, which (after initial authentication) establish only posession. At the same time, individual login attempts offer security improvements over traditional passthoughts alone, as the continuous inherence step provides an extra, ongoing validation against individual authentication attempts.

\subsection{Organic passwords}
\label{sec:orgd39940c}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org0118f24}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org50e578e}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org9e676a3}

@@

However, the original, and most active areas of research in BCI surround the creation of tools for persons with muscular disabilites \cite{Carrino2012}.

>>>>

% Created 2017-04-14 Fri 11:21
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{lmodern}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.5)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org3679620}

@@

First proposed by \cite{Thorpe2005}, ``passthoughts'' authentication allows users to 
to submit both a knowledge factor (i.e., a secret thought) and an inherence factor (i.e., the unique way that thought is expressed)
in a single step, by performing a single mental task \cite{Johnson2014}.
Since its original proposal, passthoughts has been validated in lab settings, even with 
consumer-grade devices \cite{Chuang2013b} and in-ear EEG earbuds \cite{curranpassthoughts}.
The protocol appears to be robust against impersonation attacks \cite{Johnson2014}.

@@

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life.
First, we must understand the real usability properties of passthoughts.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).
Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand 
the space of possible passthoughts,
and how passthoughts change over time (Section 4).

@@

\section{Background}
\label{sec:org7f44a3d}

@@

An ongoing problem in authentication lies in balancing strong security
(i.e., multiple factors of authentication)
with usability.
As an example, major industry players such as Google and
Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA), in which a user enters his or her password (a knowledge factor),
and subsequently receives a code on their cellphone (a posession factor).

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgf700ca0}

@@

\subsection{Passthought authentication}
\label{sec:orgd15ed7a}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org62a3127}

Even consumer-grade headsets can be uncomfortable to wear, and are awkwardly visible to outside observers. 
Earbuds present a more discreet, comfortable location for an EEG sensor, as earbuds are already commonly worn.

@@

\section{User attitudes and perceptions}
\label{sec:org36fdf32}

While past work makes passthoughts less visible with more discreet form-factors,
a large question still remains:
What sense would people make of passthoughts, as a technology, in their everyday life?
This question begs not only user-centered design studies with passthoughts itself,
but more general questions about what EEG means to people,
and what people believe EEG data can reveal about them.
Past work has established that people tend to ascribe almost magical abilities to brain-scanning devices, even subjects with specific training in the limitations of brain-scanners \cite{Ali2014a}.
Will these attitudes scare away, or attract wider adoption?
This section outlines common concerns around ``mind-reading'' machines, and how they relate to EEG and passthoughts specifically.

\subsection{Contending with mind-reading machines}
\label{sec:org0ffaa23}

@@

If people \emph{think} a certain technology measures aspects of mind, it will certainly affect the way they engage with that technology, 
whether or not it works the way they expect \cite{Ali2014a}. 
Meanwhile, if they think that a given technology does \emph{not} measure their mind, when in fact it does, users may suffer a breach of what Nissenbaum might call the ``appropriateness of the flow of information'' \cite{Doyle2011}. 
In both cases, knowing what people expect will help us anticipate their needs and concerns.

If we wish to understand what role passthought authentication \emph{could} play in day-to-day life,
we must view it both through the lens of potential privacy concerns, \emph{and} through the lens of possible opportunities for self-reflection and self-understanding. 
Of course, users' attitudes will not be fixed: they will evolve over time, as users observe the device in action, and correlate its judgments with their own lived experiences \cite{Nafus2016}.
In the next section, discuss how EEG as a sensing modality motivates questions around the meaning people may build around passthought authenticators.

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org8641b2d}

@@

\label{fig:rank}
\begin{figure*}
\centering
\includegraphics[width=.9\linewidth]{./figures/rankings.png}
\caption{``Please rank the following sensors in how likely you believe they are to reveal what a person is thinking and feeling.'' Higher bars indicate higher rank, or higher likelihood of being revealing.}
\end{figure*}

\section{Diversity and security of passthoughts}
\label{sec:orgce16208}

@@

While past work on passthoughts has achieved excellent results using recordings from different users, 
these studies do not consider a variety of different subject conditions.
For example, sitting subjects may have different patterns of neural activity from subjects who are standing, walking or exercising \cite{Thibault2016a},
let alone subjects who are under the influence of e.g. caffiene or alcohol.
Passthoughts studies must collect larger, and more diverse corpora of EEG data to examine how passthoughts change (or remain stable) throughout the dynamic contexts of daily life.

Investigating this topic could also help us understand how and why passthoughts work at all: Why are passthoughts unique, and how unique are they?
A primary question in passthoughts surrounds how large the real space of possible passthoughts might be \cite{Thorpe2005}.
While the space of possible passthoughts is potentially unlimited, we do not yet know what passthoughts we stand a reasonable chance at observing consistently over time.
A larger corpus of data might help shed light on this issue by allowing us to observe the distribution of signals that people produce over time.

A more subtle, but closely related question surrounds how passthought EEG recordings relate statistically to non-passthought EEG recordings.
In other words, we do not know how the particular passthoughts observed in past work are drawn from the distribution of EEG signals that an individual produces over the course of their day.
This blind-spot poses a possible challenge to passthought's vulnerability to dictionary-style cracking.
If an attacker has a large enough corpus of EEG readings, do some passthoughts start to look as guessable as \emph{password1234}?
By answering such questions, we could design data-driven policies for, e.g., how many retry attempts passthought authenticators should allow.

\section{Two studies on passthoughts}
\label{sec:org93415e3}

The prior two sections raise two main topics that future work could address. 
First, our limited understanding of passthoughts' usability, and user attitudes about the sensing modality present immediate questions for further development of this technology.
Second, our limited knowledge of how passthoughts shift and change over time, and around the diversity of EEG signals as our (non-medical) devices sense them,
raise questions about how frequently passthoughts would need to be calibrated, how accurate we can expect the protocol to be in different context, and how secure it might remain under threat from a motivated attacker.

This section proposes two studies on passthought authentication which, taken together, could make headway on these topics.
One study, a controlled, lab-based experiment, seeks to raise fundamental questions about how the feedback of a real-time authentication system may affect the way users perform their passthoughts.
It also begins to address certain, limited questions around the shifting nature of neural signals.
The second study, a longitudinal deployment, seeks to collect a large and diverse corpus of EEG signals, while probing people's beliefs and attitudes about EEG and brainscanning in everyday life.
Together, these studies address both long-term concerns about user attitudes and signal diversities, and also short-timescale questions about the usability and accuracy of passthoughts in realistic use scenarios.

\subsection{A real-time passthought authenticator}
\label{sec:org6270dfe}

Passthoughts promise more usable form multi-factor authentication compared to existing protocols,
as they provide both a knowledge and an inherence factor in a single-step user action.
However, no study yet has systematically evaluated passthoughts' usability.
Here, we propose a study aimed at examining passthoughts' usability in an ecologically valid context.

\subsubsection{Study protocol}
\label{sec:orgc8b5a9f}

@@

After this task, subjects could take a post-questionnaire including various usability questions.
After filling out this questionnaire, the experimenter might engage users in a brief, ten-minute semi-structured interview,
in which subjects are asked to recount their experience with the authenticator.
This interview could help gain some richer, qualitative data that traditional survey methods might fail to capture.

\subsubsection{The effect of feedback}
\label{sec:org355b8d4}

@@

Furthermore, using the data collected during this study, we could perform an offline analysis 
to test for the effect of these conditions on the actual performance of users' passthoughts.
When subjects are continuously rejected, do their passthoughts change in frustration (or in an attempt to gain access)?
We might find that passthought performance 
remains stable, regardless of what feedback subjects are shown.
Alternatively, we might find that performance changes 
when subjects are continuously rejected from their authenticator.
Alternatively, performance may change, 
even when subjects are continually accepted by their classifier.

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgbdf2dfc}

@@

By comparing the final FAR and FRR for each subject using these strategies, 
compared to the one-time calibration strategy, we could begin to get an idea as to whether
this strategy helps achieve superior performance, especially when signals change.
This analysis could also act as a harbinger for some of the possible downsides of this approach:
If a user is continually rejected, and the classifier is re-trained using those rejections as negative examples,
will the user find themselves trapped in a negative spiral of ever-decreasing authentication accuracy?

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orge0e6e8d}

@@

Unfortunately, these challenges (particularly those around shifting neural signals) also make it difficult to produce a passthought authenticator that works with any reliability in real-life contexts.
This makes a longitudinal study with a working authenticator impractical for the time being.
However, we may still perform a longitudinal study that allows us to interrogate the usability aspects around (and attitudes about) passthoughts specifically, and EEG generally.
In so doing, we may also collect a larger and more diverse corpus of passthoughts, which can be used to address the paucity of data we face today.
A technology probe or diary study \cite{Gaver1999} could help address both of these issues at the same time.

@@

\subsubsection{Study protocol}
\label{sec:org5f63529}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few checkpoints throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

Such a study would trade a large population size for a large corpus of diverse data.
This tradeoff allows us to closely investigate the diversity of EEG signals within subjects.
The diverse readings encountered in day-to-day life could help us understand how such signals change as a function of time, and/or in different psychophysical states.
At the same time, our user diaries and interviews could enable a rich, qualitative understanding of users attitudes.

\subsubsection{A more diverse corpus}
\label{sec:org9b3ef8e}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org84dd6dc}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

Leveraging the statistical clusters of EEG data generated by these algorithms, it might also be possible to generate a ``passthoughts cracker,'' capable of generating plausible passthoughts. 
Feeding these algorithms into pre-trained passthought classifiers, we can begin to generate realistic models of classifiers' resistance to cracking attempts. 
These cracking experiments could lead to defenses against cracking attempts, by enforcing retry attempt timeouts or other methods for limiting break-in risk, such that strong security guarantees can be enforced.

\subsubsection{Usability and attitudes}
\label{sec:orgec6f327}

By deploying a real sensing apparatus, be it a traditional consumer device such as the Muse \cite{Mihajlovic2015} 
or a more experimental piece of equipment such as an earbud,
and having people record EEG data in their daily life, we could learn more about the interpretative qualities of these data \cite{NafusDawn;Sherman2014}.
This study presents a dual opportunity to understand user beliefs with rich, qualitative data, 
while simultaneously collecting the large, diverse and longitudinal corpus of EEG signals necessary if we wish to stand a chance at decent authentication accuracy in the wild.

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org73c2a3d}

After the studies described above, 
we will have a much better grasp on the usability, and security properties of passthought authentication.
However, there may still be unexplored risks, challenges, and tradeoffs,
especially around user privacy.
Indeed, some of these risks are unique to the application context of biometric authentication, and to EEG as a class of biosignal. 
This section briefly reviews risks to user privacy and security that widespread passthought authentication may introduce. 
We present broad class of categories from which such risks may emerge.

\subsection{Privacy}
\label{sec:orga501979}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org8354c45}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

There is also the question of the security of data infrastructures in which EEG data might be stored.
Large data repositories are what Wolf \cite{Wolf2010} calls a ``toxic asset''; they must be maintained, 
lest the maintainer take liability for harmful fallout of poor data management.
With biosignals, as with many kinds of data, it is not clear what they might mean until they are already collected in aggregate. 
By then, it is too late to decide on an appropriate data security policy.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org6fd9d7d}

@@

\section{Further Future Directions}
\label{sec:orgaafa17b}

@@

\subsection{Continuous authentication}
\label{sec:org004177f}

After immediate challenges are overcome,
one potentially exciting possibility is that of using EEG for \emph{continuous authentication}.
Continuous authentication schemes seek to authenticate a user using ongoing streams of data or activity, sometimes by giving a probability that a person's identity is authentic \cite{Bojinov2012}.
Such schemes are a natural match for wearables, which can continuously collect and process biometric data.
A recent startup, Unify.ID, has begun to perform cross-device continuous authentication as a service \cite{UnifyID2017};
however, as a knowledge factor, it currently falls back on traditional passwords, which come with both well-known risks and annoyances to usability.

A continuous passthought authenticator could incorporate both knowledge and inherence factors (along with, optionally, the posession factor of a unique sensing device).
Subjects could perform secret passthoughts for certain unlocking actions,
while the authenticator could fall back on inherence in the base case (e.g. as an additional check on sites where the user's logged-in session would otherwise be remembered).
In theory, this strategy provides better security properties than saved sessions or cookies, 
which, after initial authentication, establish only posession. 
Individual login attempts also offer security improvements over traditional passthoughts alone, as the continuous inherence step provides an ongoing validation against individual authentication attempts.

\subsection{Organic passwords}
\label{sec:org7167579}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:orgb3c86be}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:orgf96c2fb}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org00f32ad}

@@

However, the original, and most active areas of research in BCI surround the creation of tools for persons with muscular disabilites \cite{Carrino2012}.
By collecting unstructured or semi-structured EEG data in the wild, passthought systems could help improve the development of such BCIs \cite{Grierson2011a}.
The small size of data repositories, limited mostly by the clinical trials needed to build BCIs for persons with disabilities,
has consistently frustrated attempts to improve on algorithms and protocols in this field \cite{Allison2009}.
